---
title: "initial-survival-curve"
author: "Daniel Spakowicz"
date: "5/2/2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("install.load")) {
  install.packages("install.load")
  library(install.load)
}

# CRAN packages
list.of.packages <- c("tidyverse", "survival", "survminer", "knitr", "gmodels",
                      "cgdsr")
install.load::install_load(list.of.packages)

# Define custom function for threshold calculations
avedev <- function(x) {
  sum(
    abs(
      x - mean(x)
    )
  ) / length(x)
}
```

This script is the background for checking jag1's effect on survival across all tumor types in TCGA

```{r cgds}

# Create CGDS object
mycgds = CGDS("http://www.cbioportal.org/public-portal/")

```

First I need to find all of the studies for which there are RNAseq z-scores.

```{r}
all.studies <- getCancerStudies(mycgds)
```

Filter to just TCGA.

```{r}
tcga.studies <- all.studies[grep("tcga$", all.studies$cancer_study_id),]
kable(tcga.studies[,1:2], format = "html", row.names = FALSE)
```

Just 32 studies! Great! That's a nice starting place. 

Now get the case lists

```{r, eval=FALSE}
tcga.info <- data.frame(study = tcga.studies$cancer_study_id,
                        case = NA)

for (i in 1:length(tcga.info$study)) {
  cases <- getCaseLists(mycgds, tcga.info$study[i])
  tcga.info$case[i] <- cases$case_list_id[grep("RNA Seq V2", 
                                                  cases$case_list_name)]
}
```



```{r, eval=FALSE}
# Retrieve clinical data for all mRNA cases
clinical.data <- list()
clin.cols <- c("DAYS_LAST_FOLLOWUP", "DAYS_TO_DEATH", "OS_STATUS", 
               "VITAL_STATUS", "OS_MONTHS", "")

for (s in tcga.info$case) {
  try({
    y <- getClinicalData(mycgds, s) %>%
      select(., one_of(clin.cols)) %>%
      mutate(bcr_patient_barcode = rownames(.) %>%
               str_replace("(.*)\\.01", "\\1") %>%
               str_replace_all("\\.", "-")
      )
    y$study <- s
    clinical.data[[s]] <- y
  })  
}
```

```{r jag1 expression, eval=FALSE}

expression.data <- list()

for (s in tcga.info$case) {
  try({
    # Pull the rnaseq data (v2)
    expression.data[[s]] <- getProfileData(x = mycgds,
                                  genes = c("JAG1"),
                                  geneticProfiles = s,
                                  caseList = s)
    expression.data[[s]] <- expression.data[[s]] %>%
            mutate(bcr_patient_barcode = rownames(.) %>%
               str_replace("(.*)\\.01", "\\1") %>%
               str_replace_all("\\.", "-")
      )
  })  
}
```

```{r save, eval=FALSE}
save(clinical.data, expression.data, file = "../analysis/data/raw_data/cgds_clin-expr.Rdata")
```

```{r load}
load("../analysis/data/raw_data/cgds_clin-expr.Rdata")

# Match the lists
expression.data <- expression.data[names(expression.data) %in% names(clinical.data)]
```

Now to bind the lists into dataframes that for the survival analysis.

```{r merge expression and clinical}
c.e <- list()
for (s in names(expression.data)) {
  try({
    c.e[[s]] <- inner_join(clinical.data[[s]], expression.data[[s]])
  })
}
```

Define a threshold for high expression vs other. 

```{r indicator}

ind <- lapply(c.e, function(x) x$ind <- ifelse(x$JAG1 > quantile(x$JAG1, 0.66),
                                                1, 0))

for (i in 1:length(c.e)) {
  c.e[[i]]$ind <- ind[[i]]
}


```



```{r survival p-values}

surv.pvals <- list()
for (i in names(c.e)) {
  try({
    df <- c.e[[i]] %>%
      dplyr::filter(!is.na(OS_STATUS) & OS_STATUS != '') %>%
      dplyr::filter(OS_MONTHS > 0 | !is.na(OS_MONTHS))
    
    df$surv.status <- ifelse(df$OS_STATUS == "DECEASED", 1, 0)
    
    surv.pvals[[i]] <- summary(coxph(Surv(OS_MONTHS, surv.status) ~ ind, 
                             data = df))$sctest[3]
  })
  
}

surv.pvals %>%
  unlist %>%
  data.frame %>%
  rename(pval = ".") %>%
  mutate(study = rownames(.)) %>%
  mutate(study = str_remove(study, "_tcga_rna_seq_v2_mrna.pvalue")) %>%
  ggplot() +
  geom_point(aes(x = reorder(study, pval), y = pval)) +
  coord_flip() +
  theme_bw() +
  geom_hline(yintercept = 0.05, color = "red") +
  labs(title = "P-values for surv ~ jag1 high, threshold = quantile 0.66")
```
```{r}

```

